TE_computation

### 2021-03-26 computational model/TE_computation_Simulation.Rmd
updated simulation (generating from D & N)

# Simulation 1
## Infants of the same developmental level with different word vocabularies and balances of exposure
### V = 600
```{r}
# Generating the parameters
V <- 600 # setting V = 600 for the Simulation (keep V constant)

W <- seq(0, 2*V, by = 25) # a range of total word vocabulary of a child

B <- seq(0.5, 1, by = 0.1) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
data_Simulation1 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq) %>%
  # Create other variables
  mutate(D = B*W, # Words produced in the dominant language
         N = (1-B)*W, # Words produced in non-dominant language
         TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= V)) %>%
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0)

# Visualizing Simulation 1
Simulation1 <- data_Simulation1 %>%
  # Convert to long data
  pivot_longer(c(W, D, N), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             W = "Panel A: Total vocabulary (W)", 
                             D = "Panel B: Dominant vocabulary (D)", 
                             N = "Panel C: Non-dominant vocabulary (N)")) %>%
  mutate(B = as.factor(B)) %>%
  ggplot(aes(x = number, y = TE, color = B)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5),
        text = element_text(size=12),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)"
       #, title = "Number of translation equivalent by balance in relation to different input measures "
       ) +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_Simulation1.png", Simulation1,
       width = 10, height = 5)

# Simulation 2
## Patterns of translation equivalent learning as a function of different developmental timepoints
### V = 150, 300, 450, or 600
```{r}
# V = 600
## Generating the parameters
V <- 600
W <- V # holding W constant
B <- seq(0.5, 1, by = 0.01) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)

## Generating sequence of W and B
W_seq <- W %>% 
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

## Creating the dataframe
TE_V600 <- data.frame(V, W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B)

# V = 450
## Generating the parameters
V <- 450
W <- V # holding W constant
B <- seq(0.5, 1, by = 0.01) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)

## Generating sequence of W and B
W_seq <- W %>% 
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

## Creating the dataframe
TE_V450 <- data.frame(V, W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B)

# V = 300
## Generating the parameters
V <- 300
W <- V # holding W constant
B <- seq(0.5, 1, by = 0.01) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)

## Generating sequence of W and B
W_seq <- W %>% 
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

## Creating the dataframe
TE_V300 <- data.frame(V, W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B)

# V = 150
## Generating the parameters
V <- 150
W <- V # holding W constant
B <- seq(0.5, 1, by = 0.01) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)

## Generating sequence of W and B
W_seq <- W %>% 
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

## Creating the dataframe
TE_V150 <- data.frame(V, W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B)

# Combining all the dataframes
data_Simulation2 <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
  # Create other variables
  mutate(D = B*W, # Words produced in the dominant language
         N = (1-B)*W, # Words produced in non-dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         DU = D-TE, # Unique words in dominant language
         NU = N-TE) %>% # Unique words in non-dominant language
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0)

# Visualizing Simulation 2
Simulation2 <- data_Simulation2 %>%
  mutate(bin_B = case_when(B < 0.6 ~ "0.5",
                           B >= 0.6 & B < 0.7 ~ "0.6",
                           B >= 0.7 & B < 0.8 ~ "0.7",
                           B >= 0.8 & B < 0.9 ~ "0.8",
                           TRUE ~ "0.9")) %>%
  pivot_longer(-c(V, W, B, bin_B), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("NU", "DU", "TE"))) %>%
  group_by(V, bin_B, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ V) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), 
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5),
        text = element_text(size=12),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) + 
  labs(x = "Balance", 
       y = "Number of concepts", 
       title = "Vocabulary composition across different developmental timepoints (V)")

ggsave("plot_Simulation2.png", Simulation2,
        width = 10, height = 5)

# Simulation 3
## Non-independence of dominant-language (D) and non-dominant language words (N)
```{r}
# V = 600
## Generating the parameters
V <- 600
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

## Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

## Creating the dataframe
TE_V600 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

# V = 450
## Generating the parameters
V <- 450
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

## Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

## Creating the dataframe
TE_V450 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

# V = 300
## Generating the parameters
V <- 300
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

## Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

## Creating the dataframe
TE_V300 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

# V = 150
## Generating the parameters
V <- 150
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

## Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

## Creating the dataframe
TE_V150 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

# Combining all the dataframes
data_Simulation3 <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
  # It is not possible for N > D, so keep only cases where D >= N
  filter(D >= N) %>%
  # Create TE 
  mutate(W = D+N, # Total word vocabulary of a child
         B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
         TE = D*N/V) %>% # If D and N are independent, TE = DN/V
  # Remove cases where B = NaN (those were all cases where both D and N are 0)
  filter(!is.na(B)) %>%
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0)

@@

## Visualizing the simulation
Simulation3 <- data_Simulation3 %>%
  ## Calculate where IND > 1
  mutate(TE_easier = IND1*(D*N/V)) %>%
  ## Calculate where IND < 1
  mutate(TE_harder = IND2*(D*N/V)) %>%
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0) %>%
  # Prepare dataset to plot
  pivot_longer(c(TE, TE_harder, TE_easier), 
               names_to = "Model", values_to = "number_TE") %>%
  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
  mutate(Model = recode(Model, 
                        TE = "IND = 1 (assuming independence of D and N)",
                        TE_easier = "IND = 1.5 (TEs are more easily learned)",
                        TE_harder = "IND = 0.5 (TE are less easily learned)")) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  # ggplot
  ggplot(aes(x = W, y = number_TE, color = bin_B)) +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(. ~ Model) +
  ylim(0, 600) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size=12),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  labs(x = "Total word vocabulary (W)", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_Simulation3.png", Simulation3, 
       width = 10, height = 5)

>>>>

# Simulation 1
## Infants of the same developmental level with different word vocabularies and balances of exposure
### V = 600
```{r}
V = 600

# Generating the parameters
W <- V # a range of total word vocabulary of a child
N <- seq(0, W/2, by = 10) # a range of non-dominant words
TE_W1 <- data.frame(V, W, N)

W <- V-100 # a range of total word vocabulary of a child
N <- seq(0, W/2, by = 10) # a range of non-dominant words
TE_W2 <- data.frame(V, W, N)

W <- V-100*2 # a range of total word vocabulary of a child
N <- seq(0, W/2, by = 10) # a range of non-dominant words
TE_W3 <- data.frame(V, W, N)

W <- V-100*3 # a range of total word vocabulary of a child
N <- seq(0, W/2, by = 10) # a range of non-dominant words
TE_W4 <- data.frame(V, W, N)

W <- V-100*4 # a range of total word vocabulary of a child
N <- seq(0, W/2, by = 10) # a range of non-dominant words
TE_W5 <- data.frame(V, W, N)

W <- V-100*5 # a range of total word vocabulary of a child
N <- seq(0, W/2, by = 10) # a range of non-dominant words
TE_W6 <- data.frame(V, W, N)

# Creating the dataframe
data_Simulation1 <- rbind(TE_W1, TE_W2, TE_W3, TE_W4, TE_W5, TE_W6) %>%
  # Get dominant words D
  mutate(D = W - N) %>%
  # Create other variables
  mutate(B = N/(D+N), # balance,
         TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0) %>%
  # bin B into 6 groups
  mutate(bin_B = case_when(B == 0.0 ~ "0",
                           B > 0 & B <= 0.1 ~ "0.1",
                           B > 0.1 & B <= 0.2 ~ "0.2",
                           B > 0.2 & B <= 0.3 ~ "0.3", 
                           B > 0.3 & B <= 0.4 ~ "0.4", 
                           TRUE ~ "0.5")) %>%
  arrange(bin_B)

# Visualizing Simulation 1
Simulation1 <- data_Simulation1 %>%
  # Convert to long data
  pivot_longer(c(W, D, N), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             W = "Panel 1A: Total vocabulary (W)", 
                             D = "Panel 1B: Dominant vocabulary (D)", 
                             N = "Panel 1C: Non-dominant vocabulary (N)")) %>%
  ggplot(aes(x = number, y = TE, color = bin_B, linetype = bin_B)) +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#969696", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotted", "dotdash", "dashed", "twodash", "solid")) +
  #scale_color_manual(values=c("#5e4fa2", "#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
  #scale_color_discrete(direction = -1) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 200) +
  xlim(0, 700) +
  labs(x = "Number of words", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)",
       linetype = "Balance (B)",
       title = "1) Simulated data") +
  guides(color = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

ggsave("plot_Simulation1.png", Simulation1,
       width = 10, height = 5)
```

# Simulation 2
## Patterns of translation equivalent learning as a function of different developmental timepoints
### V = 200, 400, or 600
```{r}
# V = 600
## Generating the parameters
V <- 600
W <- V # holding W constant
N <- seq(0, W/2, by = 10) # a range of non-dominant words
TE_V600 <- data.frame(V, W, N)

V <- 400
W <- V # holding W constant
N <- seq(0, W/2, by = 10) # a range of non-dominant words
TE_V400 <- data.frame(V, W, N)

V <- 200
W <- V # holding W constant
N <- seq(0, W/2, by = 10) # a range of non-dominant words
TE_V200 <- data.frame(V, W, N)

# Creating the dataframe
data_Simulation2 <- rbind(TE_V200, TE_V400, TE_V600) %>%
  # Get dominant words D
  mutate(D = W - N) %>%
  # Create other variables
  mutate(B = N/(D+N), # balance,
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         DU = D-TE, # Unique words in dominant language
         NU = N-TE) %>% # Unique words in non-dominant language
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0) %>%
  # bin B into 3 groups
  mutate(bin_B = case_when(B >= 0.35 ~ "Most Balanced",
                           B <= 0.20 ~ "Least Balanced",
                           TRUE ~ "Medium Balanced")) %>%
  mutate(bin_B  = fct_rev(bin_B)) %>%
  arrange(bin_B)

# Visualizing Simulation 2
Simulation2 <- data_Simulation2 %>%
  pivot_longer(-c(V, W, B, bin_B), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("NU", "DU", "TE"))) %>%
  group_by(V, bin_B, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ V) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
  ylim(0, 600) +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=14),
        panel.spacing.x = unit(2, "lines")) + 
  labs(x = "\nBalance (B)", 
       y = "Number of concepts (C)", 
       title = "Panel A: Simulated data")

ggsave("plot_Simulation2.png", Simulation2,
        width = 10, height = 5)
```

# Simulation 3
## Non-independence of dominant-language (D) and non-dominant language words (N)
```{r}
# V = 600
## Generating the parameters
V <- 600
W <- V
N <- seq(0, W/2, by = 10)
TE_V600 <- data.frame(V, W, N)

V <- 450
W <- V
N <- seq(0, W/2, by = 10)
TE_V450 <- data.frame(V, W, N)

V <- 300
W <- V
N <- seq(0, W/2, by = 10)
TE_V300 <- data.frame(V, W, N)

V <- 150
W <- V
N <- seq(0, W/2, by = 10)
TE_V150 <- data.frame(V, W, N)

# Creating the dataframe
data_Simulation3 <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
  # Get dominant words D
  mutate(D = W - N) %>%
  # Create other variables
  mutate(B = N/(D+N), # balance,
         TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0) %>%
  # bin B into 6 groups
  mutate(bin_B = case_when(B == 0.0 ~ "0",
                           B > 0 & B <= 0.1 ~ "0.1",
                           B > 0.1 & B <= 0.2 ~ "0.2",
                           B > 0.2 & B <= 0.3 ~ "0.3", 
                           B > 0.3 & B <= 0.4 ~ "0.4", 
                           TRUE ~ "0.5")) %>%
  arrange(bin_B)

  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0", "0.1", "0.2", "0.3", "0.4", "0.5"))) %>%

@@

## Visualizing the simulation
Simulation3 <- data_Simulation3 %>%
  ## Calculate where IND > 1
  mutate(TE_easier = IND1*(D*N/V)) %>%
  ## Calculate where IND < 1
  mutate(TE_harder = IND2*(D*N/V)) %>%
  # Round to keep whole numbers only
  mutate_at(vars(-c(B, bin_B)), round, 0) %>%
  # Prepare dataset to plot
  pivot_longer(c(TE, TE_harder, TE_easier), 
               names_to = "Model", values_to = "number_TE") %>%
  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
  mutate(Model = recode(Model, 
                        TE = "IND = 1 (assuming independence of D and N)",
                        TE_easier = "IND = 1.5 (TEs are more easily learned)",
                        TE_harder = "IND = 0.5 (TE are less easily learned)")) %>%
  # ggplot
  ggplot(aes(x = W, y = number_TE, color = bin_B, linetype = bin_B)) +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#969696", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotted", "dotdash", "dashed", "twodash", "solid")) +
  #scale_color_discrete(direction = -1) +
  theme_minimal() + 
  facet_grid(. ~ Model) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size=12),
        panel.spacing.x = unit(2.5, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  labs(x = "Total word vocabulary (W)", 
       y = "Total number of translation equivalents (TE)", 
       color = "Balance (B)",
       linetype = "Balance (B)") +
  guides(colour = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

ggsave("plot_Simulation3.png", Simulation3, 
       width = 10, height = 5)
```
